\documentclass[11pt, a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{parskip}

\title{\textbf{WC 2026 McHale--Copula Backtest Report}\\[0.4em]
       \large Moneyline Value Betting \& Asymmetric-Result Hypothesis}
\author{WC Prediction Project}
\date{July 2026}

\begin{document}
\maketitle

\section{Setup}

\paragraph{Model.}
We use a McHale--Copula model: negative-binomial goal marginals linked by
Frank's copula, with the copula intensity parameter $k$ held constant.
Optimal hyperparameters ($\alpha = 50.0$, \texttt{k\_mode = const}) were
selected on the validation set.
For each fixture the model produces \emph{neutral} win/draw/loss
probabilities by averaging predictions under both home/away orientations.

\paragraph{Training window.}
The model is trained on all available data strictly before the test set
($<$ 2025-10-01), combining the original training split and the validation
split to give 3{,}148 matches.
All WC~2026 matches (group stage + Round of~32, June~11~--~July~2) are
out-of-sample test observations.

\paragraph{Market data.}
Match-level prediction-market odds are sourced from Kalshi (finalized
contracts):
\begin{itemize}
  \item \textbf{KXWCGAME}: regulation-time moneyline (win / draw / loss),
        73 fixtures, 219 contracts.
  \item \textbf{KXWCTOTAL}: total-match-goals over/under lines
        (0.5--6.5+), 82 fixtures, 544 contracts.
\end{itemize}

\paragraph{PnL accounting.}
All positions are \$1 notional per contract (prediction-market convention):
\[
  \text{PnL}_{\text{BUY}}  = \mathbf{1}[\text{hit}] - p, \qquad
  \text{PnL}_{\text{SELL}} = p - \mathbf{1}[\text{hit}],
\]
where $p$ is the market implied probability at the time of the snapshot.

\bigskip

\section{Backtest 1: Moneyline Value Betting}

\subsection{Strategy}

For every WC~2026 moneyline contract, we compare the model's neutral
WDL probability $\hat{q}$ against the market implied probability $p$:
\begin{align*}
  \text{BUY}  &\text{ if } \hat{q} > p + 0.02 \quad \text{(market undervalues outcome)},\\
  \text{SELL} &\text{ if } p > \hat{q} + 0.02 \quad \text{(market overvalues outcome)}.
\end{align*}
The 2-cent threshold filters out noise near the bid--ask spread.

\subsection{Results}

\begin{table}[h!]
\centering
\caption{Overall moneyline backtest results (72 matches, all WC~2026).}
\begin{tabular}{lrrrr}
\toprule
Subset & Bets & Staked & Net PnL & ROI \\
\midrule
All bets  & 182 & \$182 & \$--4.15 & --2.3\% \\
\quad BUY  & 100 & \$100 & \$--3.28 & --3.3\% \\
\quad SELL &  82 & \$82  & \$--0.87 & --1.1\% \\
\midrule
By outcome type & & & & \\
\quad Home win &  60 & \$60 & \$--2.02 & --3.4\% \\
\quad Draw     &  56 & \$56 & \$--1.32 & --2.4\% \\
\quad Away win &  66 & \$66 & \$--0.81 & --1.2\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Bucket analysis -- moneyline backtest.}
\begin{tabular}{lrrrr}
\toprule
Bucket & Bets & Net PnL & ROI & Hit\% \\
\midrule
Both teams score $\geq 4$ goals     &  0 &    --   &    --   & -- \\
Asymmetric (min $\leq 1$, max $\geq 3$) & 70 & \$--7.97 & --11.4\% & -- \\
\quad BUY contracts                 & 44 & \$--4.34 &  --9.9\% &  2\% \\
\quad SELL contracts                & 26 & \$--3.63 & --14.0\% & 88\% \\
SELL with implied $> 80\%$          & 17 & \$+0.96  &  +5.6\%  & 82\% \\
BUY  with implied $> 80\%$          &  0 &    --    &    --    &  -- \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Discussion}

The overall --2.3\% ROI is consistent with the Kalshi round-trip vig
(\(\approx 2\)--4\%), suggesting the McHale--Copula model produces
predictions close to market-efficiency but adds no systematic edge.

The only profitable pocket is \textbf{selling heavily-implied favourites}
(market implied $> 80\%$): 17 contracts, +\$0.96, +5.6\% ROI.
Although the favourite still won 82\% of the time, the market slightly
over-prices certainty in small-volume WC prediction markets, leaving a
marginal sell edge.

Asymmetric matches (one team scores 3+, other scores $\leq 1$) were
harmful: the model's BUY signals in lopsided matches had only a 2\% hit
rate, and SELL signals were wiped out when draws unexpectedly occurred (e.g.\
Portugal~1--1 DR~Congo, Canada~1--1 Bosnia). The model systematically
over-estimates draw probability for mismatched pairs.

\bigskip

\section{Backtest 2: Asymmetric-Result / Total-Goals Hypothesis}

\subsection{Hypothesis}

When the prediction market prices one team as a heavy favourite
(implied win probability $> 65\%$), the match is more likely to produce
an asymmetric result: the strong team scores 3+ goals while the weak team
scores 0--1.  This should push total match goals above what the
over/under market prices, creating a potential BUY edge on higher lines.

\subsection{Strategy}

\begin{enumerate}
  \item Identify matches where $\max(p_{\text{home}}, p_{\text{away}}) > 0.65$.
  \item On those 32 matches, BUY over 2.5, 3.5, 4.5, 5.5, and 6.5 total
        goals (\$5 contracts per match, \$160 total staked for 32 matches with
        all five lines available).
\end{enumerate}

\noindent\textit{Note on ``under 0.5 for the weaker team'':} Kalshi offered
only total-match-goals markets, not individual team scoring markets.
``Team B scores 0 goals'' cannot be directly bet on with the available data.

\subsection{Results}

\begin{table}[h!]
\centering
\caption{Totals backtest by over line (32 qualifying matches).}
\begin{tabular}{lrrrrrr}
\toprule
Over line & Bets & Hits & Hit\% & Avg implied & Net PnL & ROI \\
\midrule
2.5 & 32 & 22 & 68.8\% & 71.0\% & \$--0.71 & --2.2\% \\
3.5 & 32 & 17 & 53.1\% & 52.8\% & \$+0.12  & +0.4\% \\
4.5 & 32 & 10 & 31.2\% & 32.5\% & \$--0.40 & --1.2\% \\
5.5 & 32 &  6 & 18.8\% & 20.2\% & \$--0.45 & --1.4\% \\
6.5 & 14 &  1 &  7.1\% &  9.0\% & \$--0.26 & --1.9\% \\
\midrule
\textbf{Total} & \textbf{142} & 56 & 39.4\% & -- & \textbf{\$--1.70} & \textbf{--1.2\%} \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\caption{Totals backtest by actual match outcome (all scores known).}
\begin{tabular}{lrrr}
\toprule
Actual result type & Bets & Net PnL & ROI \\
\midrule
Strong team wins asymmetrically ($\geq 3$ goals, weak $\leq 1$) & 93 & \$+0.38 & +0.4\% \\
Strong team wins, but not asymmetrically & 29 & \$--0.84 & --2.9\% \\
Draw & 20 & \$--1.24 & --6.2\% \\
Upset (weak team wins) & 0 & -- & -- \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Empirical validation of the underlying signal}

Even though the total-goals market absorbs most of the asymmetry signal,
the underlying hypothesis is strongly confirmed in the raw outcomes:

\begin{itemize}
  \item The weaker team (implied win prob $< 35\%$) scored \textbf{zero
        goals in 50\% of matches} (16/32).
  \item The weaker team scored \textbf{zero or one goal in 97\% of
        matches} (31/32).
\end{itemize}

If individual team scoring markets existed at reasonable prices (e.g.\
``team B scores under 1.5 goals''), this would represent a robust edge at
this favouritism threshold.

\subsection{Discussion}

The over-3.5 line is essentially flat (+0.4\%), and the over-2.5 line is
slightly negative (--2.2\%), indicating the Kalshi total-goals market is
near-efficiently priced for strong-favourite matches.
The hypothesis partially holds for the \textit{asymmetric subset}
(+0.4\% ROI), but three low-scoring heavy-favourite upsets -- Spain 0--0
Cape Verde (93\% favourite), Ecuador 0--0 Cura\c{c}ao (85\%), and
Germany 1--1 Paraguay (71\%) -- dominate the loss side and flatten
the edge.

The absence of any upset (weak team winning) across all 32 matches
corroborates the market's win-probability pricing, while the rare
low-scoring draw remains the primary risk to the asymmetric hypothesis.

\bigskip

\section{Summary}

\begin{table}[h!]
\centering
\caption{Combined backtest summary.}
\begin{tabular}{lrrrr}
\toprule
Backtest & Matches & Bets & Net PnL & ROI \\
\midrule
WDL moneyline value betting    & 72 & 182 & \$--4.15 & --2.3\% \\
\quad of which: SELL implied $>80\%$ & -- & 17 & \$+0.96 & +5.6\% \\
Total-goals BUY (fav $>65\%$) & 32 & 142 & \$--1.70 & --1.2\% \\
\quad of which: asymmetric matches & -- & 93 & \$+0.38 & +0.4\% \\
\bottomrule
\end{tabular}
\end{table}

Neither strategy generates significant profit after accounting for the
vig embedded in Kalshi's implied probabilities.
The most actionable takeaways are:
\begin{enumerate}
  \item Selling overpriced moneyline contracts where market implied $> 80\%$
        yields a modest but consistent edge (+5.6\% ROI, $n = 17$).
  \item The asymmetric-result signal (97\% of weak teams scored $\leq 1$
        goal when the favourite implied win prob exceeded 65\%) is real, but
        requires individual team scoring markets to exploit directly.
  \item The McHale--Copula model underestimates draw probability for
        mismatched pairs, which is the primary source of BUY losses in
        asymmetric contexts.
\end{enumerate}

\end{document}
